---
title: "Final_Project"
author: "Cathy Zhuang"
date: "2022-11-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

bba: 0 no, 1 yes
level: 1 analyst, 2 team lead, 3 director, 4 exec
housing: 0 on-campus, 1 off-campus


```{r, message = FALSE, warning = FALSE}
library(statnet)
library(kableExtra)
library(dplyr)
library(mvmeta)
library(sna)
library(keyplayer)
library(NbClust)

setwd("C:/Users/Cathy/Desktop/QTM 385 Network Analysis")
data <- as.matrix(read.csv("club_network.csv", header = T))

### Create a network object
Fnet <- network(data, directed = T)

### Import and set attributes for nodes
att <- read.csv("club_att.csv", header = T, row.names = 1 )
att$year <- 5 - (att$year - 2022)
att$bba <- att$bba - 1
att$housing <- att$housing - 1

set.vertex.attribute( Fnet, "year", att$year )
set.vertex.attribute( Fnet, "bba", att$bba)
set.vertex.attribute( Fnet, "level", att$level )
set.vertex.attribute( Fnet, "housing", att$housing )
```

Plot the network with nodes colored by “Level”.
```{r}
plot.network(Fnet, displaylabels = T, boxed.labels = FALSE, vertex.cex = 0.6, label.cex = 1, 
             vertex.col = att$level, label.col = "blue", edge.lwd = .5, edge.col = "gray50")
```
Make a table showing each actor’s degree, betweenness, closeness, and eigenvector 
centrality. Show the correlations across these measures. 
```{r}
indeg <- degree(Fnet, cmode = "indegree") # indegree centrality
outdeg <- degree(Fnet, cmode = "outdegree") # outdegree centrality

closeness <- closeness(Fnet, cmode="suminvdir") # closeness centrality
betweenness <- betweenness(Fnet) # betweenness centrality
ecent <- evcent(Fnet) # eigenvector centrality
member <- as.vector(1:length(data[,1]))



tab1 <- cbind(member, indeg, outdeg, closeness, betweenness, ecent)
colnames(tab1) <- c("Member","In-Degree","Out-Degree","Closeness", "Betweenness", "Eigenvector")
tab1 %>%  
  kbl(caption = "Centrality Measures of Each Club Member") %>%
  kable_classic_2(full_width = F)

tab2 <- cbind(indeg, outdeg, closeness, betweenness, ecent)
colnames(tab2) <- c("In-Degree","Out-Degree","Closeness", "Betweenness", "Eigenvector")
cor(tab2) %>%  
  kbl(caption = "Correlations Among Centrality Measures") %>%
  kable_classic_2(full_width = F) # show correlations among centrality measures
```
List the four members with the highest individual indegree. Calculate the total number of 
unique incoming connections they have in the network. 
```{r}
deg1 <- degree(data, cmode = "indegree") 
names(deg1) <- seq(1, 27, 1)
ids <- as.numeric(names(sort(deg1, decreasing=T)[1:4])) # Find the ids of the most central four individuals
ids %>%  
  kbl(caption = "Top 4 Members with highest individual indegree") %>%
  kable_classic_2(full_width = F)

kpcent(data,ids,type="degree", cmode="indegree", method = "max") %>%
  kbl(caption = "Total  number of unique incoming connections of the fourmembers") %>%
  kable_classic_2(full_width = F)
```
List the four members with the highest group indegree. Calculate the total number of 
unique incoming connections they have in the network. 
```{r}
groupind <- kpset(data,size=4,type="degree", cmode="indegree", method = "max")
groupind$keyplayers %>%  
  kbl(caption = "Top 4 Members with highest group indegree") %>%
  kable_classic_2(full_width = F)

groupind$centrality %>%  
  kbl(caption = "Total  number of unique incoming connections of the four members") %>%
  kable_classic_2(full_width = F)
```

Run a linear regression to predict the in-degree centrality using the members' attributes. 

```{r}
regression <- cbind(indeg, att)
regression1 <- lm(indeg ~ year + bba + level + housing, data = regression)

regression1$coefficients %>% 
  kbl( caption = "Linear Regression of In-Degree Centrality Using Member Attributes") %>%
  kable_classic_2(full_width = F)
```
List all the cliques with 3 or more members. 
```{r}
clique <- clique.census(Fnet)

# See clique counts
cliquecounts <- clique$clique.count[,1]
cliquecounts

# From the clique counts we can see that there are cliques with 4 and 5 members (cliques with 3 or more members)

tab3 <- clique$cliques[c(3,4,5)]

tab5 <- data.frame(clique$cliques[3])
colnames(tab5) <- c(1:length(tab5))
tab5  %>% 
  kbl(caption = "Cliques with 3 Members") %>%
  kable_classic_2(full_width = F)

tab6 <- data.frame(clique$cliques[4])
colnames(tab6) <- c(1:length(tab6))
tab6 %>% 
  kbl(caption = "Cliques with 4 Members") %>%
  kable_classic_2(full_width = F)

tab7 <- data.frame(clique$cliques[5])
colnames(tab7) <- c(1:length(tab7))
tab7  %>% 
  kbl(caption = "Cliques with 5 Members") %>%
  kable_classic_2(full_width = F)

```
Calculate the mutuality, density, centralization, and transitivity of the network. Provide boxplots that 
compare these statistics in the actual network with those in randomized networks.
```{r}
# mutuality
mutuality <- grecip(Fnet, measure = "edgewise")
print(paste0("The mutuality is ", mutuality))

# density
density <- gden(Fnet)
print(paste0("The density is ", density))

# centralization
cent <- centralization(Fnet, degree)
print(paste0("The centralization is ", cent))

# transitivity
transitivity <- gtrans(Fnet)
print(paste0("The transitivity is ", transitivity))


##### Randomization test for network structures


### Randomize the network and get the statistics/estimates
## Fix the number of ties each actor sent 
## But randomize to whom the ties are directed to
## Vectors to store statistics from randomized networks
mut <- NULL
trans <- NULL
dens <- NULL
centr <- NULL

# Get the adjacency matrix
FMat <- as.matrix(Fnet)
K <- nrow(FMat)
# number of randomization
M <- 200

for (i in 1:M) {
  S <- sample(K)
  RMat <- FMat[, S]
  RNet <- network(RMat)
  
  mut <- append(mut, grecip(RNet, measure = "edgewise"))
  dens <- append(dens, gden(RNet))
  centr <- append(centr, centralization(RNet, degree))
  trans <- append(trans, gtrans(RNet))
}

boxplot(mut, ylim=c(0,1.2), main = "Mutuality of Network Compared to Randomized Network"  )
abline(h=density, col = "Red")


boxplot(dens, ylim=c(0,1.2), main = "Density of Network Compared to Randomized Network"  )
abline(h=density, col = "Red")


boxplot(centr, ylim=c(0,1.2), main = "Centralization of Network Compared to Randomized Network"  )
abline(h=cent, col = "Red")


boxplot(trans, ylim=c(0,0.5), main = "Transitivity of Network Compared to Randomized Network" )
abline(h=transitivity, col = "Red")

```

Use “brokerage” and “level” to calculate a brokerage index (i.e., the total number of times 
an actor serving as a broker). Which three actors have the highest brokerage power? How 
much does your brokerage index correlate with the betweenness centrality? 

```{r}
sb <- brokerage(Fnet, att$level)
# Observed brokerage scores, by vertex. t column shows the total roles.
sb$raw.nli %>%
  kbl(caption = "Brokerage Index") %>%
  kable_classic_2(full_width = F)

tab8 <- cbind(betweenness, sb$raw.nli)
cor(tab8) %>%
  kbl(caption = "Correlation of Brokerage Index with Betweenness Centrality") %>%
  kable_classic_2(full_width = F)
```


What does structural equivalence mean here? Show the structural equivalence matrix. 
Based on the structural equivalence matrix, present a hierarchical clustering of the actors. 

In  this case, two members are considered structurally equivalent if they are connected to other  members in a similar way.

```{r}
se <- sedist (Fnet, method = "euclidean") # use euclidean distance
se[upper.tri(se)] <- ""
se <- as.data.frame(se)
se <- cbind(a = c(1:nrow(se)), se)
se <- as.data.frame(apply(se, 2, as.numeric))
se <- se %>%
  mutate(across(where(is.numeric), round, digits=1))
colnames(se) <- c(0:(ncol(se)-1))


se  %>%
  kbl(caption = "Structural Equivalence Matrix") %>%
  kable_classic_2(full_width = F)

# Hierarchical clustering of network positions
eq <- equiv.clust (Fnet, method="euclidean") # use euclidean distance

plot(eq, main="Hierarchical clustering of network positions", xlab = "", sub="")
```

Divide the actors into three blocks based on structural equivalence. Present the density 
matrix of the three blocks. 
```{r, fig.width=20,fig.height=10}
# This approach maximizes structural equivalence within blocks.
# divide into blocks, let's do 3 blocks
block <- blockmodel(Fnet, eq, k = 3 )


# block modeling
plot(block, displaylabels = F, xlab="", ylab="", main="Blockgrouping", sub="")

block$block.model %>%   
  kbl(caption = "Density Matrix of 3 Blocks") %>%
  kable_classic_2(full_width = F)
  
# gives you density table, also tells you which manager belongs to which blocks
```
# ERGM

```{r}
## Covariate Effects
# nodeicov: receiver effects for continuous variables
# nodeocov: sender effects for continuous variables
# nodecov: connection effects for continuous variables - don't distinguish between directions
# nodeifactor, nodeofactor, and nodefactor for categorical variables

## Homophily
# absdiff: difference for continuous variables
# nodematch: being the same for binary/categorical variables

# year, bba, level, housing


model1 <- ergm(Fnet ~ nodeicov("year") + nodeocov("year") + nodefactor("bba") + nodeicov("level") + nodeocov("level") + nodefactor("housing")
               + absdiff("year") + nodematch("bba") + absdiff("level") + nodematch("housing") + edges)

summary(model1)

model1summary<- as.data.frame(summary(model1)$coef)[,c(1,2,4,5)]
rownames(model1summary) <- c("Year (Receiver)", "Year (Sender)", "BBA", "Level (Sender)", "Level (Receiver)", "Housing", "Year Difference", "Same Major (BBA)", "Different Levels", "Same Housing", "Edges")

table1<- model1summary %>%   
  kbl(caption = "Summary of Model 1") %>%
  kable_classic_2(full_width = F) %>%
  pack_rows(index = c("Main Effects" = 6, "Homophily" = 4, "Endogenous Network  Formation" = 1))

table1
```

Had to force MPLE and assume naive logistic regression, AIC came out better though. Choose model 2

```{r}
# gw stands for geometrically weighted
# gwidegree is the term for describing preferential attachment - the tendency for a person to send a tie to a popular other (person who gets lot of ties)
# gwodegree is differential sociality - you prefer to send ties to a person who already sends out a lot of ties
# gwesp is transitivity
# gwdsp is a term to control two-path 

model2 <- ergm(Fnet ~ nodeicov("year") + nodeocov("year") + nodefactor("bba") + nodeicov("level") + nodeocov("level") + nodefactor("housing")
               + absdiff("year") + nodematch("bba") + absdiff("level") + nodematch("housing") + edges + mutual
               + gwidegree(0.1, fixed = TRUE) + gwodegree(0.1, fixed = TRUE) + gwesp(0.1, fixed = TRUE) + gwdsp(0.1, fixed = TRUE), estimate = "MPLE", 
               control = control.ergm(MCMLE.maxit = 300, MCMC.burnin=10000, MCMC.interval=200, seed = 178 ))

summary(model2)

model2summary <- as.data.frame(summary(model2)$coef)[,c(1,2,4,5)] 
rownames(model2summary) <- c("Year (Receiver)", "Year (Sender)", "BBA", "Level (Sender)", "Level (Receiver)", "Housing", 
                             "Year Difference", "Same Major (BBA)", "Different Levels", "Same Housing", 
                             "Edges", "Mutuality", "Preferential Attachment", "Differential Sociality", "Transitivity", "Two-Path")
 

table2<- model2summary %>%   
  kbl(caption = "Summary of Model 2") %>%
  kable_classic_2(full_width = F) %>%
  pack_rows(index = c("Main Effects" = 6, "Homophily" = 4, "Endogenous Network  Formation" = 6))

table2
```

Goodness of fit
```{r, fig.width=40,fig.height=20}
gof2 <- gof(model2)
par(mfrow=c(1,5))
plot(gof2)
```











